Stack Safety as a Security Property

### 2020-06-29 main.tex
Write-up about the run-from-state version

The {\it split} operation \(\mathit{split}(f,T)\) takes a trace \(T\)
and a predicate on elements \(f\) and separates \(T\) into a prefix
and suffix at the first element on which \(f\) holds:

\judgment{\(f\ t\)}
         {\(\mathit{split}(f,tT) = t, tT\)}
%  
\judgmenttwo{\(\neg f\ t\)}{\(\mathit{split}(f,T) = T_1,T_2\)}
            {\(\mathit{split}(f,tT) = tT_1,T_2\)}
%
Note that the last element of the prefix and the first element of the
suffix coincide. As we will mostly be using $\mathit{split}$ to
partition a trace to a callee and a caller portion, this choice will
allow us to preserve local state encapsulation invariants during
the transition between the two.

The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
$\mathit{split}$: it takes two traces and links them by replacing the
last element of the first (if it has one) with the first element of
the second

\judgment{} {\(t \cdot T = T\)}
\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
\leo{Lemma/theorem: split/join = id?}
\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}

Finally, we will also want to take a prefix of the trace from an
initial state up through the first state where some condition holds.
This can easily be implemented using $\mathit{split}$, but as we're
going to be making heavy use of it we're going to provide special
notation for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
\(\machT\) up to \(f\)''):

\judgmenttwo{\(\mach \hookrightarrow \machT\)}
            {\(\mathit{split}(\lambda (\mach',\obs).f(\mach'),\machT)
              = (\machT_1,\machT_2)\)}
            {\(\mach \hookrightarrow \machT_1 | f\)}
\noindent
And similarly for traces with policies:

  \judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
              {\(\mathit{split}(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT_1,\MPT_2\)}
         {\(\mpstatename \hookrightarrow \MPT_1 | f\)}

@@

>>>>

The {\it prefUpTo} operation \(\PUT(\Trace,f)\) takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

\judgment{\(f\ \traceelem\)}
         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
%
\judgment{}
         {\(\PUT(f,\traceelem) = \traceelem\)}
%
\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. Conversely, if the predicate never
holds, \(\PUT\) is the identity function.

% The {\it join} operation, written \(T_1 \cdot T_2\) is the reverse of
%$\mathit{split}$: it takes two traces and links them by replacing the
%last element of the first (if it has one) with the first element of
%the second

%\judgment{} {\(t \cdot T = T\)}
%\judgment{\(T_1 \cdot T_2 = T\)}{\(tT_1 \cdot T_2 = tT\)}
%\leo{Lemma/theorem: split/join = id?}
%\leo{TODO: FORMATTING: Remove force line break/par at end of judgments}

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
running \(\mach\) up to \(f\)''):

\judgmenttwo{\(\mach \hookrightarrow \machT\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT)
              = \machT'\)}
            {\(\mach \hookrightarrow \machT' | f\)}
\noindent
And similarly for traces with policies:

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
         {\(\mpstatename \hookrightarrow \MPT' | f\)}

@@

\sna{Version for stepping from a state}
\definition{Confidential Execution}

  Stack confidentiality, below, is a property of a system that, from a call to
  a return, each trace obeys a local confidentiality property,
  {\it confidential execution}.

  We will say that a state \(\mpstate{\mach}{\pol}\) satisfies confidential
  execution with regard to a contour \(\contour\) and an ending condition
  \(R\) that is a predicate on machine states if, for all variant states
  \(\mach' \approx_\contour \mach\), the induced traces from \(\mach\) and
  \(\mach'\) behave the same across three cases. Let
  \(\mpstate{\mach}{\pol} \hookrightarrow \MPT | R\) and
  \(\mach' \hookrightarrow \machT | R\).

  If \(\MPT\) ends with a state \(\mach_r\) and \(R\ \mach_r\), then \(\machT\)
  ends with a state \(\mach_r'\) such that \(R\ \mach_r'\). \(\MPT\) and
  \(\machT\) are observationally equivalent, and for each component
  \(\component\) such that \(\contour(\component) = \HIGHSEC\),
  \(\mach_r(\component) = \mach_r'(\component)\).

  If \(\MPT\) is infinite, then \(\machT\) is infinite, and they are
  observationally equivalent.

  If \(\MPT\) ends before \(R\) ever holds, meaning that it halted due to a
  policy violation, then it is a strict observational prefix of \(\machT\),
  which must be infinite.
